---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
df <- read.csv(unz('activity.zip', 'activity.csv'))
noNA <- na.omit(df)
head(df)
```


## What is mean total number of steps taken per day?
```{r}
# Remove NA
total <- aggregate(noNA$steps, by=list(noNA$date), FUN=sum)
names(total) <- c('date', 'steps')
total$date <- as.Date(as.character(total$date), '%Y-%m-%d')
```

The total number of steps per day are:
```{r}
total
```

In histogram form:
```{r}
hist(total$steps, main='Total number of steps per day', xlab='Steps per day')
```

The mean total number of steps per day is:
```{r}
mean(total$steps)
```

The median total number of steps per day is:
```{r}
median(total$steps)
```


## What is the average daily activity pattern?

```{r}
# assuming we can still ignore NAs
averages <- aggregate(noNA$steps, by=list(noNA$interval), FUN=mean)
names(averages) <- c('interval', 'steps')
plot(averages, type='l', main='Average number of steps per interval')
```

The interval with the maximum average number of steps is:
```{r}
averages$interval[which.max(averages$steps)]
```

## Inputing missing values
The number of rows with missing values is:
```{r}
sum(is.na(df))
```

Fill missing values with the average amount of steps in an interval
```{r}
library(dplyr)
filled <- df %>% mutate_at(vars(steps), ~ifelse(is.na(.), mean(., na.rm=TRUE), .))
```

In histogram form:
```{r}
total <- aggregate(filled$steps, by=list(filled$date), FUN=sum)
names(total) <- c('date', 'steps')
hist(total$steps, main='Total number of steps per day', xlab='Steps per day')
```

The mean total number of steps per day is:
```{r}
mean(total$steps)
```

The median total number of steps per day is:
```{r}
median(total$steps)
```

We can see that the average and median is very similar but we can see that the frequency of days where between 10000 and 15000 steps were made increased.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# Add new factor
filled$date <- as.Date(as.character(filled$date), '%Y-%m-%d')
weekdaysStr <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
filled$day <- factor(weekdays(filled$date) %in% weekdaysStr, labels=c('weekend', 'weekday'))

# Compute weekday averages
week <- subset(filled, day == 'weekday')
avgWeek <- aggregate(week$steps, by=list(week$interval), FUN=mean)
names(avgWeek) <- c('interval', 'steps')

# Compute weekend averages
weekend <- subset(filled, day == 'weekend')
avgWeekend <- aggregate(weekend$steps, by=list(weekend$interval), FUN=mean)
names(avgWeekend) <- c('interval', 'steps')

# Build the plot
par(mfrow=c(2,1))
plot(avgWeek, type='l', ylab='Number of steps')
plot(avgWeekend, type='l')


```

